---
title: "HW6"
author: "Raphael Nash"
date: "3/6/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning=FALSE, message=FALSE}
library(Rmpfr)
library(textmineR)
```

#Problem Set 1

1. When you roll a fair die 3 times, how many possible outcomes are there
```{r}
6^3
```


2. What is the probability of getting a sum total of 3 when you roll a die two times?

```{r}
2/36
```

3. Assume a room of 25 strangers. What is the probability that two of them have thesame birthday? Assume that all birthdays are equally likely and equal to 1/365 each. What happens to this probability when there are 50 people in the room?

**Probability for number of people is 25**
```{r}

roundMpfr(1 - (factorialMpfr(365) / factorialMpfr((365 - 25))) / (365^25), 15)

```

**Probability for number of people is 50**
```{r}

roundMpfr(1 - (factorialMpfr(365) / factorialMpfr((365 - 50)))/(365^50), 15) 

```


# Problem Set 2
Sometimes you cannot compute the probability of an outcome by measuring the sample
space and examining the symmetries of the underlying physical phenomenon, as you could
do when you rolled die or picked a card from a shuffled deck. You have to estimate
probabilities by other means. For instance, when you have to compute the probability of
various english words, it is not possible to do it by examination of the sample space as it is
too large. You have to resort to empirical techniques to get a good enough estimate. One
such approach would be to take a large corpus of documents and from those documents,
count the number of occurrences of a particular character or word and then base your
estimate on that.
Write a program to take a document in English and print out the estimated probabilities
for each of the words that occur in that document. Your program should take in a le
containing a large document and write out the probabilities of each of the words that
appear in that document. Please remove all punctuation (quotes, commas, hyphens etc)
and convert the words to lower case before you perform your calculations.
Extend your program to calculate the probability of two words occurring adjacent to
each other. It should take in a document, and two words (say the and for) and compute
the probability of each of the words occurring in the document and the joint probability
of both of them occurring together. The order of the two words is not important.
Use the accompanying document for your testing purposes. Compare your probabilities
of various words with the Time Magazine corpus: http://corpus.byu.edu/time/

<br><br>
**Probability of individual Words:**
```{r  warning=FALSE, message=FALSE }

x <- readLines("assign6.sample.txt")
x <- x[ x != ""]
x <- iconv(x,to = "utf-8-mac")


dtm <- CreateDtm(doc_vec = x,
                 ngram_window = c(1, 1),
                 lower = TRUE,
                 remove_punctuation = TRUE,
                 remove_numbers = TRUE ,
                 stopword_vec = c() )

word_probability_df <- as.data.frame (colSums(as.matrix(dtm) )) 
word_probability_df$word <- rownames(word_probability_df)
colnames(word_probability_df) <- c("probability", "word")

word_probability_df <-
  word_probability_df[order(word_probability_df$probability, decreasing = TRUE), ]
word_probability_df$probability <- 
  word_probability_df$probability / sum(colSums(as.matrix(dtm) ))
rownames(word_probability_df) <- c()
head(word_probability_df, 20)
```
<br><br>
**Probability of bi grams:**
```{r  warning=FALSE, message=FALSE}

dtm <- CreateDtm(doc_vec = x,
                ngram_window = c(2, 2),
                stopword_vec = c()  , 
                lower = TRUE,
                remove_punctuation = TRUE,
                remove_numbers = TRUE)

bigram_probability_df <- as.data.frame (colSums(as.matrix(dtm) )) 
bigram_probability_df$word <- rownames(bigram_probability_df)
colnames(bigram_probability_df) <- c("probability", "bigram")

bigram_probability_df <- 
  bigram_probability_df[order(bigram_probability_df$probability, decreasing = TRUE), ]
bigram_probability_df$probability <- 
  bigram_probability_df$probability / sum(colSums(as.matrix(dtm) ))
rownames(bigram_probability_df) <- c()

head(bigram_probability_df, 20) 

```

I could not understand how to get probabilities out of the website: http://corpus.byu.edu/time/.  I could only get word counts by decade. 
