---
title: "Rnash-Assing11.rmd"
author: "Raphael Nash"
date: "4/10/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

#Problem 1
**Using R's lm function, perform regression analysis and measure the significance of the
independent variables for the following two data sets. In the first case, you are evaluating
the statement that we hear that Maximum Heart Rate of a person is related to their age
by the following equation:

MaxHR = 220 - Age**

```{r}
hr_df= data.frame(
  Age = c(18,23,25,35,65,54,34,56,72,19,23,42,18,39,37),
  MaxHR = c(202,186,187,180,156,169,174,172,153,199,193,174,198,183,178)
)

hr_model <- lm (hr_df$MaxHR~ hr_df$Age)
summary(hr_model)
```


**What is the resulting equation?**

$$y = `r hr_model$coefficients[['hr_df$Age']] ` \times Age - `r hr_model$coefficients[['(Intercept)']]` $$
```{r}
hr_model$coefficients[[2]]
hr_model$coefficients[[1]]
```

**What is the signiffcance level?** 

` r hr_model_stats$coefficients[2,4]`

**Please also plot the fitted relationship between Max HR and Age.**

```{r}
ggplot(data=hr_df, aes(y=MaxHR, x= Age)) +geom_point() + stat_smooth(method = "lm", col = "red")

```

# Problem 2

**Using the Auto data set from Assignment 5 (also attached here) perform a Linear Re-
gression analysis using mpg as the dependent variable and the other 4 (displacement, horse-
power, weight, acceleration) as independent variables. What is the final linear regression fit
equation?**

**Perform the following by a) using the all the data, b) Using a sample of 40 data points:**

```{r}
mpg_data <- read.table("https://raw.githubusercontent.com/RaphaelNash/CUNY-DATA-605-CompMath/master/HW11/auto-mpg.data")
names(mpg_data) <- c('displacement', 'horsepower', 'weight', 'acceleration', 'mpg')
```

```{r}

set.seed(1)
mpg_sample_data <- mpg_data[sample(nrow(mpg_data), 40), ]
mpg_sample_model <- lm(mpg ~ ., mpg_sample_data)
mpg_sample_model_ci <- confint(mpg_sample_model)
mpg_sample_model_summary <- summary(mpg_sample_model)
mpg_sample_model_coef <- mpg_sample_model_summary$coefficients


mpg_sample_model_sig <- mpg_sample_model_coef[, "Pr(>|t|)"]
mpg_sample_model_summary
```

```{r}

mpg_model <- lm(mpg ~ ., mpg_data)
mpg_model_ci <- confint(mpg_model)
mpg_model_summary <- summary(mpg_model)
mpg_model_coef <- mpg_model_summary$coefficients
mpg_model_sig <- mpg_model_coef[, "Pr(>|t|)"]
mpg_model_summary

```

**What is the final linear regression fit?**



Sample Model: mpg = (`r mpg_sample_model_coef[['displacement', 1]]` )  displacement + (`r mpg_sample_model_coef[['horsepower',1]] `)  horsepower + (`r mpg_sample_model_coef[['weight',1]]`)  weight + (`r mpg_sample_model_coef[['acceleration',1]] `)  acceleration +  (`r mpg_sample_model_coef[['(Intercept)',1]] `)

Full Model:  mpg = (`r mpg_model_coef[['displacement', 1]]` )  displacement + (`r mpg_model_coef[['horsepower',1]] `)  horsepower + (`r mpg_sample_model_coef[['weight',1]]`)  weight + (`r mpg_model_coef[['acceleration',1]] `)  acceleration +  (`r mpg_model_coef[['(Intercept)',1]] `)


**Which of the 4 independent variables have a significant impact on mpg? **


```{r}

sig <- cbind(mpg_sample_model_sig, mpg_model_sig) 
sig

```


Sample:   weight only

Full Model:  displacement, weight, acceleration

**What are the standard errors on each of the coeffcients?**


Sample Model:
```{r}
standard_errors <- rbind ( mpg_sample_model_coef[,'Std. Error'], mpg_model_coef[,'Std. Error']) 
rownames(standard_errors) <- c('sample_model', 'full_model')
standard_errors
```

**measure the 95% confidence intervals.**

```{r}
colnames( mpg_sample_model_ci) <-c('low_sample_model', 'high_sample_model')
colnames( mpg_model_ci) <-c('low_full_model', 'high_full_model')
ci <- cbind(mpg_sample_model_ci, mpg_model_ci) 

```


**Conclusions**
1) The size of the confidence intervals is smaller on the full model 
2) The standard error is smaller on the full model
3) The number of signifigant factors is higher on the full model

