---
title: "Untitled"
author: "Raphael Nash"
date: "2/1/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem Set 1

**(1) Calculate the dont product u.v where u = [0.5;0.5] and v = [3;-4]**


```{r}
u <- matrix( c(.5,.5) )  
v <- matrix ( c( 3,-4))
 
t(u) %*%(v)
```

** What are the lengths of u and v? Please note that the mathematical notion of the
length of a vector is not the same as a computer science deffnition.**


```{r}

length_u <- sqrt(t(u) %*% u )
length_u

lenght_v <- sqrt(t(v) %*% v )
lenght_v
```

**(2) What is the linear combination 3u-2v?**
```{r }
3*u-2*v
```

#Problem Set 2

**Set up a system of equations with 3 variables and 3 constraints and solve for x. Please
write a function in R that will take two variables (matrix A & constraint vector b) and
solve using elimination. Your function should produce the right answer for the system of
equations for any 3-variable, 3-equation system. You don't have to worry about degenerate
cases and can safely assume that the function will only be tested with a system of equations
that has a solution. Please note that you do have to worry about zero pivots, though. Please
note that you should not use the built-in function solve to solve this system or use matrix
inverses. The approach that you should employ is to construct an Upper Triangular Matrix
and then back-substitute to get the solution. Alternatively, you can augment the matrix
A with vector b and jointly apply the Gauss Jordan elimination procedure.** 


Test with: 

$\begin{pmatrix}
1 & 1 & 3 \\
2 & -1 & 5 \\
-1 & -2 & 4 
\end{pmatrix}
\begin{pmatrix}
X_1\\ 
X_2\\
X_3
\end{pmatrix}= 
\begin{pmatrix}
1 \\
2 \\
6 
\end{pmatrix}$

Should have the soultion of x= [-1.55, -0.32, 0.95]



```{r}
linearEqSolve <- function ( A,C ) {
  
  #---------------Create upper Triangle Matrix
  aug <- cbind(A,C)
  for ( cur_row in 2: nrow(aug)) {
    for( cur_col in 1: (cur_row -1 ))  {
      denom <- aug[cur_col,cur_col]
      num <- aug [cur_row ,cur_col]
      mult <- num / denom 
      aug[cur_row,] <-  aug[cur_row,] - (aug[cur_col,] * mult ) 

    }
  
  }
  u <- aug[,1:ncol(aug)-1]
  v <- aug[,ncol(aug)]
  
  print("Augmented Upper Triangle Matrix:") 
  print(aug)


  x <- matrix( , nrow =   nrow(u), ncol = 1 )
  cur_row <-0 
  cur_col <-0
  
  #----------Back Solve  
  for (cur_row in ncol(u) : 1) {
    cur_val <-  v[cur_row]
    for( cur_col  in   ncol(u) : cur_row ) {
      offset <- cur_col- cur_row 
      if ( offset == 0 ) {  
        cur_val <- cur_val  / u[cur_row, cur_row]
        x[cur_row] <- cur_val
      }  else {
        cur_val <- cur_val - u[cur_row, cur_row+offset] * x[cur_row+offset]
      } 
    }
  }
  
x
}


```

```{r}
A <- matrix (c (1,1,3,2,-1,5,-1,-2,4), nrow = 3, ncol=3 , byrow = TRUE)
C <- matrix (c (1,2,6), nrow = 3, ncol=1 , byrow = TRUE)
x<- linearEqSolve(A,C)
x

```