---
title: "HW8"
author: "Raphael Nash"
date: "3/23/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Set 1

Your colleague either commutes by train or by the bus. 20 days of the month,
she takes the train and the remaining 10 days she takes the bus. If she takes the
train, she reaches work on time with a probability of 0.9. If she takes the bus, she
frequently gets stuck in traffic and reaches work on time with a probability of 0.5.
Given that she was on time today, what is the probability that she took the bus to
work today?

```{r}

p_take_bus <- 10/30
p_take_train <- 20/30
p_train_on_time  <- .9
p_bus_on_time <- .5

p_took_bus_today <- (p_take_bus * p_bus_on_time  ) / 
  ((p_bus_on_time * p_take_bus ) + (p_train_on_time *p_take_train  ))

p_took_bus_today

```


## Problem Set 2

Bayes network


```{r message= FALSE, warning=FALSE}
library("gRain")
```


Setup Bayes Newwork:
```{r}

no_yes <- c("no", "yes")
low_high <- c("low","high")
 
difficulty <- cptable(~difficulty, values=c(0.3, 0.7), levels = no_yes )
 
intelligence <- cptable(~intelligence, values=c(0.8, 0.2), levels=low_high)
 
sat_intelligence <- cptable(~sat|intelligence, values=c(0.9, 0.1, 0.2, 0.8), levels=low_high)
 
grade_intelligence_difficulty <- 
     cptable(~grade|intelligence:difficulty,
             values=c(0.6, 0.4, 0.01, 0.99, 0.8, 0.2, 0.1, 0.9 ),levels=low_high)

letter_grade <- cptable(~letter|grade,values=c(0.9, 0.1, 0.05, 0.95),levels=low_high)

p_list <- compileCPT(list(difficulty, intelligence,
                          sat_intelligence, grade_intelligence_difficulty, letter_grade))

network <- grain(p_list)
```

View Bayes Network:
```{r}

plot(network)

```
In the Grade Network that we looked at in the notes, what happens to the probability of
Diffculty of Course when you present the evidence that the received recommendation let-
ter was good?
```{r}

querygrain(setEvidence(network, "letter", "high"), "difficulty", type="marginal")

```
In addition, now present the evidence that both SAT scores were good and
the letter of recommendation was good, What is the probability of the Diffculty of Course
now?

```{r}

querygrain(setEvidence(network,  c("sat", "letter"), c("high", "high")), "difficulty")

```