---
title: "CUNY DATA 605 (CompMath) - Assign2"
author: "Raphael Nash"
date: "2/1/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Problem Set 1

**1) Show that $A^T \times A \ne A \times A^T$**


**Proof** 
$A = \begin{pmatrix}
a & b  \\
c & d 
\end{pmatrix}$

$A^T = \begin{pmatrix}
a & c  \\
b & d 
\end{pmatrix}$

$A^T \times A = \begin{pmatrix}
a^2 + c^2 & ab+cd  \\
ba+dc & b^2+d^2 
\end{pmatrix}$

$A \times A^T = \begin{pmatrix}
a^2 + b^2 & ac+bd  \\
ca+db & c^2+d^2 
\end{pmatrix}$


**Demonstration  $A^T \times A \ne A \times A^T$ **

```{r}
A <- matrix (c (1,2,3,4), nrow = 2, ncol=2 , byrow = TRUE)
A_Transpose <- t(A)

A
A_Transpose

```


$A \times A^T$


```{r}
A %*% A_Transpose
```

$A_T \times A$


```{r}
 A_Transpose %*% A 
```


**2)For a special type of square matrix A, we get $A^T \times A = AA^T$ . Under what conditions
could this be true?**

This is true if the matrix is symetrical.

$A = \begin{pmatrix}
a & c  \\
c & a 
\end{pmatrix}$

$A^T = \begin{pmatrix}
a & c  \\
c & a 
\end{pmatrix}$

Therfore:

$A^T \times A = A \times A^T = A^T$



#Problem Set 2

```{r}


LU_Factor <- function  (x) { 
u <- x

elimMatrices <- list()
i<-0
  for ( cur_row in 2: nrow(u)) {
   
    for( cur_col in 1: (cur_row -1 ))  {
       i <- i + 1 
      denom <- u[cur_col,cur_col]
      num <- u [cur_row ,cur_col]
      mult <- num / denom 
      
      elim <- diag ( nrow(u))
      elim[ cur_row, cur_col] <- -1 * mult

      u <- elim %*% u 
      
      elimMatrices[[i]] <- elim

    }
  
  }

  L <- diag(nrow(u))
  for (i in 1:length(elimMatrices))  {
    L <- L %*% solve(elimMatrices[[i]])
  }
  
  LU <- list ( L = L , U = u )
  
  LU
  
  }

A <- matrix (c (1,1,3,2,-1,5,-1,-2,4), nrow = 3, ncol=3 , byrow = TRUE)
A

LU <- LU_Factor(A)

LU

LU[["L"]] %*% LU[["U"]] 

```


