---
title: "RNash_Assign3"
author: "Raphael Nash"
date: "2/14/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(pracma)

```
#Problem Set 1 

**What is the rank of Matrix A?**

$A = \begin{pmatrix}
1 & 2 & 3 & 4  \\
-1 & 0 & 1 & 3\\
0 & 1 & -2 & 1\\
5 & 4 & -2 & -3
\end{pmatrix}$

```{r}
A <- matrix ( c(1,2,3,4,-1, 0,1,3,0 ,1,-2,1,5,4,-2,-3 ), byrow  = TRUE, ncol = 4 )
A
rref(A)
```

The rank of Matrix A is 4 

**(2) Given an mxn matrix where m > n, what can be the maximum rank? The mini-
mum rank, assuming that the matrix is non-zero?** 

The minimum rank is 1 

The maximimum ranks is m (where m is the number of rows)

**(3) What is the rank of matrix B?**

$A = \begin{pmatrix}
1 & 2 & 1\\
3 & 6 & 3 \\
2 & 4 & 2
\end{pmatrix}$

```{r}

B <- matrix (c ( 1,2,1,3,6,3,2,4,2), byrow = TRUE, ncol = 3)
B
rref(B)
```
The rank of B is 1

#Problem Set 2

**Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your
work. You'll need to write out the characteristic polynomial and show your solution.**

$A = \begin{pmatrix}
1 & 2 & 3\\
0 & 4 & 5 \\
0 & 0 & 6
\end{pmatrix}$


$\lambda I_3 - A =  \begin{pmatrix}
\lambda -1 & -2 & -3\\
0 & \lambda - 4 & -5 \\
0 & 0 & \lambda -6
\end{pmatrix}$


$det ( \lambda I_3 -A ) = 0$

$(\lambda-6)(\lambda-4)(\lambda-1)-(-2)(0-3)(0)$

$(\lambda-6)(\lambda-4)(\lambda-1) = 0$

$\lambda = 6$ or $\lambda=4$ or $\lambda = 1$

characteristic polynomial = $x^3 - 9x^2 +24x - 16$


Vector for $\lambda=3$

$N( \begin{pmatrix}
1 -1 & -2 & -3\\
0 & 1- 4 & -5 \\
0 & 0 & 1 -6
\end{pmatrix}) = 
N(\begin{pmatrix}
0 & -2 & -3\\
0 & -3 & -5 \\
0 & 0 & -5
\end{pmatrix}) = \begin{pmatrix}
0 & 1 & 0\\
0 & 0 & 1 \\
0 & 0 & 0
\end{pmatrix}$

$\begin{pmatrix}
0 & 1 & 0\\
0 & 0 & 1 \\
0 & 0 & 0
\end{pmatrix}
\begin{pmatrix}
v_1\\
v_2\\
v_3\\\end{pmatrix} = \begin{pmatrix}
0 \\
0\\
0\\\end{pmatrix}$

$\begin{pmatrix}
1\\
0\\
0\\\end{pmatrix}$



Vector for $\lambda=6$

$N( \begin{pmatrix}
6 -1 & -2 & -3\\
0 & 6- 4 & -5 \\
0 & 0 & 6 -6
\end{pmatrix}) = 
N(\begin{pmatrix}
5 & -2 & -3\\
0 & 2 & -5 \\
0 & 0 & 0
\end{pmatrix}) =\begin{pmatrix}
1 & 0 & -1.6\\
0 & 1 & -2.5 \\
0 & 0 & 0.0
\end{pmatrix}$

$\begin{pmatrix}
1 & 0 & -1.6\\
0 & 1 & -2.5 \\
0 & 0 & 0.0
\end{pmatrix}
\begin{pmatrix}
v_1\\
v_2\\
v_3\\\end{pmatrix} = 
\begin{pmatrix}
0 \\
0\\
0\\\end{pmatrix}$




$\begin{pmatrix}
-1.6\\
-2.5\\
0\\\end{pmatrix}$




Vector for $\lambda=4$

$N( \begin{pmatrix}
4 -1 & -2 & -3\\
0 & 4- 4 & -5 \\
0 & 0 & 4 -6
\end{pmatrix}) = 
N(\begin{pmatrix}
3 & -2 & -3\\
0 & 0 & -5 \\
0 & 0 & -2
\end{pmatrix}) =\begin{pmatrix}
1 & -0.6667 & 0\\
0 & 0 & 1 \\
0 & 0 & 0
\end{pmatrix}$

$\begin{pmatrix}
1 & -0.6667 & 0\\
0 & 0 & 1 \\
0 & 0 & 0
\end{pmatrix}
\begin{pmatrix}
v_1\\
v_2\\
v_3\\\end{pmatrix} = 
\begin{pmatrix}
0 \\
0\\
0\\\end{pmatrix}$




$\begin{pmatrix}
0.66667\\
1\\
0\\\end{pmatrix}$

















