---
title: "Assign10.rmd"
author: "Raphael Nash"
date: "4/4/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('matrixcalc')
library('igraph')
```

#Playing with PageRank
You'll verify for yourself that PageRank works by performing calculations on a small
universe of web pages.
Let's use the 6 page universe that we had in the course notes. For this directed graph,
perform the following calculations in R.


Form the A matrix. 

```{r}
row1 <- c(0, 1/2, 1/2, 0, 0, 0)
row2 <- c(1/6, 1/6, 1/6, 1/6, 1/6, 1/6)
row3 <- c(1/3, 1/3, 0, 0, 1/3, 0)
row4 <- c(0, 0, 0, 0, 1/2, 1/2)
row5 <- c(0, 0, 0, 1/2, 0, 1/2)
row6 <- c(0, 0, 0, 1, 0, 0)
A <- matrix(c(row1,row2, row3, row4, row5, row6), 6)
A

colSums(A)
```

Then, introduce decay and form the B matrix as we did in
the course notes.
```{r}
d <- 0.85

B <- 0.85*A + (0.15/6)
B

```


Start with a uniform rank vector r and perform power iterations on B till conver-
gence. That is, compute the solution $r = B^n \times r$. Attempt this for a sufficiently
large n so that r actually converges.

```{r}

r <- rep(1/nrow(A), nrow(A)) 
r


matrix.power(B, 10) %*% r 
matrix.power(B, 20) %*% r 
matrix.power(B, 30) %*% r 
pr_power <- matrix.power(B, 40) %*% r 
 

```
After 30 iterations the results stabalize


Compute the eigen-decomposition of B and verify that you indeed get an eigenvalue
of 1 as the largest eigenvalue and that its corresponding eigenvector is the same
vector that you obtained in the previous power iteration method. Further, this
eigenvector has all positive entries and it sums to 1.

```{r, warning=FALSE, messages=FALSE}
decomp <- eigen(B)
pr_eigen <- as.numeric(decomp$vectors[,which.max(decomp$values)]) 
pr_eigen <- (1/sum(pr_eigen))*pr_eigen
pr_eigen
 
```

Use the graph package in R and its page.rank method to compute the Page Rank
of the graph as given in A. Note that you don't need to apply decay. The package
starts with a connected graph and applies decay internally. Verify that you do get
the same PageRank vector as the two approaches above.
```{r}

ga <- graph.adjacency(t(A), weighted=TRUE, mode='directed')
ceb <- cluster_edge_betweenness(ga)
plot(ceb, ga) 
pr_igraph_package <- page.rank(ga)$vecto
pr_igraph_package
 
```

Verifcation

Verify all vectors sum to 1
```{r}
c(colSums(pr_power) ,sum(pr_eigen) ,sum(pr_igraph_package)) == 1
```

Verify all Page Ranks are the Same (atleast first 6 digits)
```{r warning=FALSE, error=FALSE}
round(pr_power[,1],6) == round(pr_eigen,6)
round(pr_power[,1],6) == round(pr_igraph_package,6)
 
```