---
title: "HW7"
author: "Raphael Nash"
date: "3/14/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Problem Set 1


Function for Mean (non-streaming)

$\bar{x} = \frac{1}{n}\sum\limits_{i=1}^{i}(X_i)$
```{r}

mean_nonstreaming <- function( x ) {
  accumulator <- 0 
  for( val in x) {
    accumulator <- accumulator + val
   
  }
  
  accumulator / length( x )

}

```


```{r}

nums <- c( 45,66,77,1,99,5,3,5)
c( mean( nums ) , mean_nonstreaming( nums ) ) 

```


Standard Deviation (non-streaming)

$\sigma = \sqrt{ \frac{1}{n}\sum\limits_{i=1}^{i}(x = \bar{x})^2}$
```{r}

sd_nonstreaming <- function( x ) {
  n <- length( x )
  mu <- mean_nonstreaming( x ) 

  acc <- 0
  for ( val in x ) {
 
    acc <- acc +  ((val-mu) ^2) 
  
  }
  
  sqrt( acc * (1/(n-1) )) # using n-1 becasue r sd function uses n-1 
                          # see the r help for details 
  
}

```

```{r}

c(sd(nums), sd_nonstreaming(nums))

```

Mean Streaming

To calulate the streaming mean, you need to store the sum of all values and the number of values in a global variable.  Then with each number you add you increment your number of observations and add the observation to the global accumulator. 
```{r}

mean_streaming <- function (x, init = TRUE) {
  
  if ( init == TRUE) {
    streaming_mean_n <<- 0 
    streaming_mean_sum <<- 0
  }
  streaming_mean_n <<- length(x) + streaming_mean_n
  
  for (val in x ) {
    streaming_mean_sum <<- val + streaming_mean_sum
  }
  
  streaming_mean_sum / streaming_mean_n
  
}

```


```{r}
rolling_mean <- mean_streaming(1, TRUE)
c( rolling_mean, mean(c(1))) 
rolling_mean <- mean_streaming(c(1,2), FALSE)
c(rolling_mean, mean(c(1,1,2))) 

```


Standard Deviation Streaming

If you re-write the standard deviation formula as:

$\sigma = \sqrt{ \frac{ \Sigma{x^2}}{n-1}-(\frac{\Sigma {x}}{n-1})^2}$

Then, if you store in globals $\Sigma{x^2}$ , $\Sigma {x}$ and $n$, you can recalculate the stadard deviation everytime you add observation to the stream.  
```{r}
streaming_sd <- function (x, init = TRUE) {
  
  if (init == TRUE) {
    streaming_sd_sum <<- 0
    streaming_sd_sumsq <<- 0
    streaming_sd_n <<- 0
  }
  
  streaming_sd_n <<- streaming_sd_n + length(x)
  streaming_sd_sum <<- streaming_sd_sum + sum(x)
  streaming_sd_sumsq <<- streaming_sd_sumsq + sum( x^2)
  
  sqrt((streaming_sd_sumsq - ((streaming_sd_sum^2) / streaming_sd_n ))*(1/(streaming_sd_n-1)))
  
}
```



```{r}

rolling_sd <- streaming_sd(c(9,10, 11), TRUE) 
c( rolling_sd, c(sd(c(9,10,11))))
rolling_sd <- streaming_sd(c(15), FALSE) 
c( rolling_sd, c(sd(c(9,10,11, 15))))


```



