---
title: "HW5"
author: "Raphael Nash"
date: "2/26/2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("matlib")   
```

# Problem Set1

**In this problem set we'll work out some properties of the least squares solution that we
reviewed in the weekly readings. Consider the unsolvable system Ax = b as given below:**

$$
\begin{pmatrix}
1 & 0\\
1 & 1 \\
1 & 3 \\
1 & 4 
\end{pmatrix} 
\begin{pmatrix} 
x_1 \\
x_2
\end{pmatrix} = 
\begin{pmatrix}
0 \\
8 \\
8 \\
20 \\
\end{pmatrix}
$$


```{r}
A <- matrix( c(1,0,1,1,1,3,1,4), ncol = 2, byrow = TRUE)
b <-  matrix ( c ( 0, 8, 8, 20), ncol = 1 )
b
```

**Write R markdown script to compute $A^TA$ and $A^Tb$.**


$A^TA$ 
```{r}
At_A <- t(A) %*% A
At_A
```

$A^TB$
```{r}
At_b  <- t(A) %*% b
At_b
```

Solve for $\hat x$ in R using the above two computed matrices
```{r}

x_hat <- solve(At_A, At_b)
x_hat
```


What is the Squared error of this solution?
$||A \hat x - b ||^2$
```{r}
sum((A %*% x_hat - b)^2)
```
Instead of b = [0; 8; 8; 20], start with p = [1; 5; 13; 17] and find the exact solution (i.e. show that this system is solvable as all equations are consistent with each other. This should result in an error vector e = 0).
```{r}
p <-  matrix ( c ( 1, 5, 13, 17), ncol = 1 )
p
```
$A^Tp$
```{r}
At_p  <- t(A) %*% p
x_hat <- solve(At_A, At_b)
x_hat
sum((A %*% x_hat - p)^2)
```

Show that the error e = b - p = [-1;3;-5;3]
```{r}
e <- b - p
e
```

Show that the error e is orthogonal to p and to each of the columns of A.
```{r}
t(e) %*% p 
```
```{r}
t(e) %*% A[,1]
```

```{r}
t(e) %*% A[,2]
```


If you take the dot product of 2 vectors and they are orthogonal, then the result should be 0.

# Problem Set 2 

Consider the modified auto-mpg data (obtained from the UC Irvine Machine Learning dataset). This dataset contains 5 columns: displacement, horsepower, weight, acceleration, mpg. We are going to model mpg as a function of the other four variables.

Write an R markdown script that takes in the auto-mpg data, extracts an A matrix from the first 4 columns and b vector from the fifth (mpg) column. Using the least squares approach, your code should compute the best fitting solution. That is, find the best fitting equation that expresses mpg in terms of the other 4 variables. Finally, calculate the fitting error between the predicted mpg of your model and the actual mpg. Your script should be able to load in the 5 column data set, extract A and b, and perform the rest of the calculations. Please have adequate comments in your code to make it easy to follow your
work.

```{r}
mpg_data <- read.table('auto-mpg.data')
A<- as.matrix(mpg_data[,c(1:4)])
head(A)
b <- as.matrix(mpg_data[,5])
head(b)
```


$A^TA$ 
```{r}
At_A <- t(A) %*% A
At_A
```

$A^TB$
```{r}
At_b  <- t(A) %*% b
At_b
```

Solve for $\hat x$ in R using the above two computed matrices
```{r}

x_hat <- solve(At_A, At_b)
x_hat
```

```{r}
#function from homwork 1 
linearEqSolve <- function ( A,C ) {
  
  #---------------Create upper Triangle Matrix
  aug <- cbind(A,C)
  for ( cur_row in 2: nrow(aug)) {
    for( cur_col in 1: (cur_row -1 ))  {
      denom <- aug[cur_col,cur_col]
      num <- aug [cur_row ,cur_col]
      mult <- num / denom 
      aug[cur_row,] <-  aug[cur_row,] - (aug[cur_col,] * mult ) 

    }
  
  }
  u <- aug[,1:ncol(aug)-1]
  v <- aug[,ncol(aug)]
  
  x <- matrix( , nrow =   nrow(u), ncol = 1 )
  cur_row <-0 
  cur_col <-0
  
  #----------Back Solve  
  for (cur_row in ncol(u) : 1) {
    cur_val <-  v[cur_row]
    for( cur_col  in   ncol(u) : cur_row ) {
      offset <- cur_col- cur_row 
      if ( offset == 0 ) {  
        cur_val <- cur_val  / u[cur_row, cur_row]
        x[cur_row] <- cur_val
      }  else {
        cur_val <- cur_val - u[cur_row, cur_row+offset] * x[cur_row+offset]
      } 
    }
  }
  
x
}

```


```{r}
solution <- linearEqSolve(At_A, At_b)
```

Linear Eq for gas mileage:

`r round(solution[1,],2)` col1 * `r round(solution[2,],2)` col2 + `r round(solution[3,],2)` col3 + `r round(solution[4,],2)` col4 = miles per gallon

